---
title: "Annual Nominal Fish Catches-googleVis Motion Charts "
author: "OmaymaS"
date: "June 21, 2016"
output: 
  html_document: 
    keep_md: yes
    self_contained: no
    toc: yes
---
# Overview
This script presents the annual TLW (tonnes live weight) catches of fish and shellfish in the Northeast Atlantic region.
The data is cleaned and prepared to plot two googleVis interactive charts. The first is for the annual TLW per country and the second is for the annual TLW per Species. *Further modifications are to b  added later*.


```{r echo=F,warning=F,message=F}
#libraries
library(dplyr)
library(reshape2)
library(googleVis)
op<-options(gvis.plot.tag = "chart")
```
## Loading and Cleaning Data

```{r}
#read data
dat<-read.csv("../data/fish_catches.csv",header = T,stringsAsFactors = F)
country_codes<-read.csv("../data/country_codes.csv",header = T,stringsAsFactors = F)

#remove NA columns
dat<-select(dat,Species:X2006)

#remove x from years 
names(dat)[5:13]<-sub('.', '', names(dat)[5:13])

#melt the data frame to long format
datm<-melt(dat,id=names(dat)[1:4],
           variable.name = "Year",
           value.name = "Quantity")

#Na.omit
datm<-na.omit(datm)

#change year from factor to numeric
datm$Year<-as.numeric(as.character(datm$Year))

#extract country and code from the country_codes dataframe
cc<-country_codes[,1:2]

#merge the main data and country codes 
datm<-left_join(datm,cc,by=c("Country"="Code"))

#rename columns 
names(datm)[names(datm)=="Country"]="Code"
names(datm)[names(datm)=="Description"]="Country"
```
Now, after melting the original dataframe and the adding the country name we have the following dataframe:
```{r echo=F}
head(datm)
```

## gvis Motion Charts

### Calculate Annual Sum per Country and Species

Here we would like to plot the total amount of fish catch grouped by country then by species. So first we will create the dataframes to be used.
```{r}
#Total tonnage by country
datm.country<-datm %>%
        group_by(Year,Country) %>%
        summarise(TLW=sum(Quantity)) %>%
        arrange(Country)

#Total tonnage by species
datm.species<-datm %>%
        group_by(Year,Species) %>%
        summarise(TLW=sum(Quantity)) %>%
        arrange(Species)
```

### Create gvisMotionCharts

Then we will create two gvisMotionCharts and plot them. But we will set its state first to show the line plot and use unique colors by default.

```{r results='asis', tidy=FALSE}
myState<-'
{"yLambda":1,"showTrails":false,"xAxisOption":"_TIME",
"xZoomedDataMin":1136073600000,"orderedByY":false,"playDuration":15000,
"nonSelectedAlpha":0.4,"xZoomedIn":false,"xLambda":1,
"iconKeySettings":[],"xZoomedDataMax":1388534400000,
"dimensions":{"iconDimensions":["dim0"]},
"sizeOption":"_UNISIZE","yZoomedIn":false,
"uniColorForNonSelected":false,"time":"2014","yZoomedDataMax":10000000,
"duration":{"timeUnit":"Y","multiplier":1},
"yAxisOption":"2","yZoomedDataMin":0,"orderedByX":false,
"iconType":"LINE",
"colorOption":"_UNIQUE_COLOR",
}
'

#create the country chart
CountryChart = gvisMotionChart(datm.country, "Country", "Year",
                    options = list(showChartButtons=F,
                                   state=myState))
#create the species chart
SpeciesChart= gvisMotionChart(datm.species, "Species", "Year",
                    options = list(showChartButtons=F, state=myState))
#Plot the 2 charts 
BothCharts<-gvisMerge(CountryChart,SpeciesChart)
#remove caption
BothCharts$html$caption=""
plot(BothCharts)
